---
title: "Smithersball"
author: "Dakota Olson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preperation
```{r load data, message=FALSE}
library(tidyverse)
pitching_91_93 <- read_csv('data/FG_Player_Pitching_1991_1993.csv')
batting_91_93 <- read_csv('data/FG_Player_Batting_1991_1993.csv')
```

## Pitcher Analysis
```{r clean pitching_91_93}
# Remove pitchers who didn't pitch a complete game between 1991 and 1993.
pitching_91_93 <- subset(pitching_91_93, CG != 0)
pitching_91_93
```

### Modify fWAR and RA9-WAR to be Rate Statistics
```{r calculate fWAR/200 IP and RA9-WAR/200 IP}
pitching_91_93 %>%
  mutate('fWAR/200 IP' = round(200 * fWAR / IP, digits=1)) -> pitching_91_93
  
pitching_91_93 %>%
  mutate('RA9-WAR/200 IP' = round(200 * pitching_91_93$'RA9-WAR' / IP, digits=1)) -> pitching_91_93

summary(pitching_91_93)
```

### Top ERA- Performers
```{r ERA-}
pitching_91_93 %>%
  arrange(pitching_91_93$'ERA-') %>%
  head(n=10)
```

### Top FIP- Performers
```{r FIP-}
pitching_91_93 %>%
  arrange(pitching_91_93$'FIP-') %>%
  head(n=10)
```

### Top fWAR/200 IP Performers
```{r fWAR/200 IP}
pitching_91_93 %>%
  arrange(-pitching_91_93$'fWAR/200 IP') %>%
  head(n=10)
```

### Top RA9-WAR/200 IP Performers
```{r RA9-WAR/200 IP}
pitching_91_93 %>%
  arrange(-pitching_91_93$'RA9-WAR/200 IP') %>%
  head(n=10)
```

### Pitcher Comparisons
```{r}
top_pitchers <- c('Jose Rijo', 'Kevin Appier', 'Greg Maddux', 'Bill Swift', 'Juan Guzman')
subset(pitching_91_93, Name %in% top_pitchers)
```

```{r Maddux vs Rijo vs Appier}
top_pitchers2 <- c('Jose Rijo', 'Kevin Appier', 'Greg Maddux')
subset(pitching_91_93, Name %in% top_pitchers2)
```

```{r Rijo vs Clemens}
top_pitchers3 <- c('Jose Rijo', 'Roger Clemens')
subset(pitching_91_93, Name %in% top_pitchers3)
```

## Batter Analysis
### Estimated wOBA Skill (Regression Towards Mean) 
```{r}
mean_wOBA <- sum(batting_91_93$PA * batting_91_93$wOBA / sum(batting_91_93$PA))
mean_wOBA
```

```{r}
var_wOBA <- sum((batting_91_93$PA/sum(batting_91_93$PA)) * (batting_91_93$wOBA - mean_wOBA) ^ 2)
var_wOBA
```

```{r ewOBA and fWAR/600 PA}
batting_91_93 %>%
  mutate('ewOBA' = round(mean_wOBA + (PA / (PA + (0.25 / var_wOBA))) * (wOBA - mean_wOBA), digits=3)) -> batting_91_93

batting_91_93 %>%
  mutate('fWAR/600 PA' = round(600 * fWAR / PA, digits=1)) -> batting_91_93
# Note: fWAR includes defensive production so fWAR per PA isn't necessarily the best metric, but this is just for simple analysis.

summary(batting_91_93)
```

```{r}
library(skimr)
skim(batting_91_93)
```

```{r}
wOBA_comp <- batting_91_93[, c('Name', 'PA', 'wOBA', 'ewOBA')]

wOBA_comp <- unique(wOBA_comp)

wOBA_comp %>%
  mutate('delta wOBA' = ewOBA - wOBA) -> wOBA_comp
  
wOBA_comp %>%
  arrange(wOBA_comp$'delta wOBA') %>%
  head(n=10)

wOBA_comp %>%
  arrange(-wOBA_comp$'delta wOBA') %>%
  head(n=10)
```

### Top ewOBA Performers
```{r ewOBA}
batting_91_93 %>%
  arrange(-ewOBA) %>%
  head(n=30)
```

### Top wRC+ Performers
```{r wRC+}
batting_91_93 %>%
  arrange(-batting_91_93$'wRC+') %>%
  head(n=30)
```

### Top fWAR/600 PA Performers
```{r fWAR/600 PA}
batting_91_93 %>%
  arrange(-batting_91_93$'fWAR/600 PA') %>%
  head(n=60)
```

### Top RF - The Danny Tartabull Emergency
```{r}
pos_of <- c('LF', 'CF', 'RF', 'OF')
best_of <- subset(batting_91_93, Position %in% pos_of) 

best_of %>%
  arrange(-best_of$'fWAR/600 PA') %>%
  head(n=20)
```

```{r}
rf_options <- c('Danny Tartabull', 'Juan Gonzalez', 'Kevin Mitchell', 'Bobby Bonilla', 'Albert Belle', 'Tim Salmon', 'David Justice', 'Shane Mack')
rf_tib <- subset(batting_91_93, Name %in% rf_options & Position %in% c('LF', 'RF'))

rf_tib %>%
  arrange(-rf_tib$'fWAR/600 PA') %>%
  head(n=11)
```

```{r}
rf_options2 <- c('Danny Tartabull', 'David Justice')
subset(batting_91_93, Name %in% rf_options2)
```

### Top CF
```{r}
pos_cf <- c('CF')
best_cf <- subset(batting_91_93, Position %in% pos_cf) 

best_cf %>%
  arrange(-best_cf$'ewOBA') %>%
  head(n=10)
```

### Top IF
```{r}
pos_if <- c('2B', '3B', 'SS')
best_if <- subset(batting_91_93, Position %in% pos_if) 

best_if %>%
  arrange(-best_if$'ewOBA') %>%
  head(n=10)
```

### Top C
```{r}
pos_c <- c('C')
best_c <- subset(batting_91_93, Position %in% pos_c) 

best_c %>%
  arrange(-best_c$'ewOBA') %>%
  head(n=10)
```

### The Team
```{r}
squad_names <- c('Chris Hoiles', 'Frank Thomas', 'Lou Whitaker', 'Barry Larkin', 'Edgar Martinez', 'Barry Bonds', 'Lenny Dykstra', 'Danny Tartabull')
the_squad <- subset(batting_91_93, Name %in% squad_names)

the_squad
```

```{r}
sum(the_squad$PA * the_squad$wOBA / sum(the_squad$PA))
sum(the_squad$PA * the_squad$ewOBA / sum(the_squad$PA))
sum(the_squad$PA * the_squad$'wRC+' / sum(the_squad$PA))
sum(the_squad$PA * the_squad$'fWAR/600 PA' / sum(the_squad$PA))
```

### Smithers Team
```{r}
the_smiths <- c('Mike Scioscia', 'Don Mattingly', 'Steve Sax', 'Ozzie Smith', 'Wade Boggs', 'Jose Canseco', 'Ken Griffey Jr.', 'Darryl Strawberry')
smith_squad <- subset(batting_91_93, Name %in% the_smiths)

smith_squad
```
